Are mental properties supervenient on brain properties?

### 2011-07-27 nature_sci_rep/submission2/supervenience.tex
organized up to including simulation in main text

\noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be the space of all possible mental properties, including all possible thoughts, memories beliefs, etc.
% respects (in whatever sense Davidson meant \cite{Davidson70}). 
%For example, $m$ might indicate a person's intelligence, psychological state, current thought, gender identity, etc.   
Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be the set of all possible brain properties,
% physical respects (which we assume to mean all possible brain properties, 
including the position and momentum of all subatomic particles within the skull.  
% For example, $b$ might denote the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$. 
Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.  This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix 1 for more details and some examples).  To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.

@@

\subsection*{Testing for statistical supervenience} % (fold)
\label{sub:subsection_name}

@@

Ideally, we might consider the following hypothesis test: $H_0: L^*>0$ and $H_A: L^*=0$.  Unfortunately, because the null hypothesis is not closed, the alternate hypothesis lies on the boundary, so the $p$-value is always equal to unity \cite{Bickel2000}.  To proceed, therefore, we introduce a relaxed notion of supervenience:
\begin{defi}
\label{def2}
Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
\end{defi}
\noindent Given this relaxation, consider the problem of testing for $\eps$-supervenience:
\begin{align*}
	H_0^{\eps}: L^* \geq \eps \\
	H_A^{\eps}: L^* < \eps
\end{align*}
Let $\mh{n}= n' \hL^{n'}_{F}(g_n)$ be the test statistic.  The distribution of $\mh{n}$ is available under the least favorable null distribution to the alternate hypothesis.  For the above hypothesis test,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', \eps)= \sum_{k \in [\mh{n}]_0}$Binomial$(k; n'; \eps)$, where $[\mh{n}]_0=\{0,1,\ldots, \mh{n}\}$.  We therefore reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.   The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a viable statistical test of supervenience, given a specified $\eps$ and $\alpha$.

% Because Binomial$(k; n', 0)=1$ for all $n'$, the $p$-value is always equal to unity.    Thus, we can \emph{never} reject the above null hypothesis, no matter how much data we obtain.

% Although the above theorem is of potential theoretical interest, the argument relies on knowing the typically unavailable $L^*$, rendering it useless pragmatically.  However,  $g^*$ could be estimated from data, and its estimate could be utilized to conduct the following statistical test of supervenience:
% % Let $g_n$ be a classifier induced from some training data, $\mc{T}_n$, 
% \begin{align*} 
% 	H_0:& L^* > 0 \\
% 	H_A:& L^*=0
% \end{align*}

% \subsection*{$\eps$-supervenience admits hypothesis testing} % (fold)
% \label{sub:hypothesis_testing}
% 
% We therefore consider a relaxed notion of supervenience:
% 
% 
% The $p$-value of the above composite alternate hypothesis is given by the binomial distribution under the least favorable distribution, $\mathbb{B}(\mh{n}; n', p_{H_A}=\eps)<1$.

\subsection*{The power and limits of an ideal test of supervenience} % (fold)
\label{sub:uc}

Ideally, our test is \emph{consistent}; that is, as $n$ increases, the power of a test (the probability of a rejecting the null when it is indeed false)  goes to unity. For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (point mass at zero), then whenever $\alpha >0$, power goes to unity.  For the above $\eps$-supervenience statistical test, if $g_n \conv g^*$ as $n \conv \infty$, then $\hL^{n'}_F(g_n) \conv L^*$ as $n,n' \conv \infty$.  Thus, if $L^* < \eps$, the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \eps) \conv \delta_0$.

@@

The above considerations indicate that to construct a statistical test with guaranteed limiting unity power, a universally consistent classifier (one for which $\hL^{n'}_F(g_n) \conv L_F(g^*)$ as $n,n' \conv \infty$) is required. Below, we show that under a very general mind-brain model, a consistent classifier is readily available.

@@

\begin{thex}
Let the physical property under consideration be brain connectivity structure, so $b$ is a brain-graph (``connectome'') with vertices representing neurons (or collections thereof) and edges representing synapses (or collections thereof). Further let $\mB$, the brain observation space, be the collection of all graphs on a finite number of vertices, and $\mc{M}$, the mental property observation space, be finite. Now, imagine collecting very large amounts of very accurate identically and independently sampled  brain-graph data and the associated mental property indicators. A $k_n$-nearest neighbor classifier using a Frobenius norm is universally consistent (see Methods for details). Therefore, %Theorem \ref{thm1} applies and 
the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind-brain property pair, if indeed $\varepsilon$-supervenience holds. This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes, as well as unlabeled graphs (see \cite{VP11unlabeled} for details). Furthermore, the proof holds for other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite (again, see Methods for details).  
\end{thex}

>>>>

\noindent Let $\mc{M}=\{m_1, m_2, \ldots\}$ be the space of all possible mental properties, including all possible thoughts, memories beliefs, etc.
Similarly, let $\mc{B}=\{b_1,b_2,\ldots\}$ be the set of all possible brain properties,
including the position and momentum of all subatomic particles within the skull.  
Given these definitions, Davidson's conjecture may be concisely and formally stated thusly:  $m \neq m' \implies b \neq b'$, where $(m,b), (m',b') \in \mc{M} \times \mc{B}$ are mind-brain pairs.  This mind-brain supervenience relation does not imply an injective relation, a causal relation, or an identity relation (see Appendix 1 for more details and some examples).  To facilitate both statistical analysis and empirical investigation, we convert this supervenience relation from a logical to a probabilistic relation.

@@

\subsection*{The existence of a consistent statistical test for supervenience} % (fold)
\label{sub:subsection_name}

@@

Ideally, we might consider the following hypothesis test: $H_0: L^*>0$ and $H_A: L^*=0$.  Unfortunately, because the null hypothesis is not closed, the alternate hypothesis lies on the boundary, so the $p$-value is always equal to unity \cite{Bickel2000}.  To proceed, therefore, we introduce a relaxed notion of supervenience:
\begin{defi}
\label{def2}
Given $\varepsilon > 0$, $\mM$ is said to $\varepsilon$-\textit{supervene} on $\mB$ for distribution $F=F_{MB}$, denoted $\MeB$, if and only if $L_{F}(g^*) < \varepsilon$.
\end{defi}
\noindent Given this relaxation, consider the problem of testing for $\eps$-supervenience:
\begin{align*}
	H_0^{\eps}: L^* \geq \eps \\
	H_A^{\eps}: L^* < \eps
\end{align*}
Let $\mh{n}= n' \hL^{n'}_{F}(g_n)$ be the test statistic.  The distribution of $\mh{n}$ is available under the least favorable null distribution to the alternate hypothesis.  For the above hypothesis test,  the $p$-value is given by the binomial cumulative distribution function, $\mathbb{B}(\mh{n}; n', \eps)= \sum_{k \in [\mh{n}]_0}$Binomial$(k; n'; \eps)$, where $[\mh{n}]_0=\{0,1,\ldots, \mh{n}\}$.  We therefore reject whenever this $p$-value is less than $\alpha$; rejection implies that we are $100(1-\alpha$)\% confident that $\MeB$.

A test is \emph{consistent} whenever its power  (the probability of a rejecting the null when it is indeed false)  goes to unity as $n$ increases. For any statistical test, if the $p$-value converges in distribution to $\delta_0$ (point mass at zero), then whenever $\alpha >0$, power goes to unity.  For the above $\eps$-supervenience statistical test, if $g_n \conv g^*$ as $n \conv \infty$, then $\hL^{n'}_F(g_n) \conv L^*$ as $n,n' \conv \infty$.  Thus, if $L^* < \eps$, the $p$-value converges; that is, $\mathbb{B}(\mh{n}; n', \eps) \conv \delta_0$.

The definition of $\eps$-supervenience therefore admits, for the first time to our knowledge, a viable statistical test of supervenience, given a specified $\eps$ and $\alpha$. Moreover, this test is consistent whenever $g_n$ converges to the Bayes classifier, $g^*$.

@@

\subsection*{The availability of a consistent statistical test for supervenience} % (fold)
\label{sub:uc}

The above considerations indicate the existence of a consistent test for supervenience whenever the classifier used is consistent.  
% that to construct a statistical test with guaranteed limiting unity power, a universally consistent classifier (one for which $\hL^{n'}_F(g_n) \conv L_F(g^*)$ as $n,n' \conv \infty$) is required. 
To actually implement such a test, one must be able to (i) measure mind/brain pairs and (ii) have a consistent classifier $g_n$.  Unfortunately, we do not know how to measure the entirety of one's brain, much less one's mind. Moreover, even if we did, we do not know the joint distribution of mind/brain pairs, $F_{MB}$.  We therefore must restrict our interest to a mental property/brain property pair.  
A mental property might be a person's intelligence, psychological state, current thought, gender identity, etc.  A brain property might be the number of cells in a person's brain at some time $t$, or the collection of spike trains of all neurons in the brain during some time period $t$ to $t'$.  Regardless of the details of the specifications of the mental property and the brain property, given such specifications, one can assume a model, $\mc{F}$.  We would like a classifier, $g_n$, that is guaranteed to be consistent, no matter which of the possible distributions $F_{MB} \in \mc{F}$ is the true distribution.  A classifier with such a property is called an \emph{universally consistent classifier}.  Below, we show that under a very general mind-brain model $\mc{F}$, an universally consistent classifier is readily available.

@@

\begin{thex}
Let the physical property under consideration be brain connectivity structure, so $b$ is a brain-graph (``connectome'') with vertices representing neurons (or collections thereof) and edges representing synapses (or collections thereof). Further let $\mB$, the brain observation space, be the collection of all graphs on a given finite number of vertices, and $\mc{M}$, the mental property observation space, be finite. Now, imagine collecting very large amounts of very accurate identically and independently sampled  brain-graph data and the associated mental property indicators from $F_{MB}$. A $k_n$-nearest neighbor classifier using a Frobenius norm is universally consistent (see Methods for details). Therefore, %Theorem \ref{thm1} applies and 
the existence of a universally consistent classifier guarantees that eventually (in $n,n'$) we will be able to conclude $\MeB$ for this mind-brain property pair, if indeed $\varepsilon$-supervenience holds. This logic holds for directed graphs or multigraphs or hypergraphs with discrete edge weights and vertex attributes, as well as unlabeled graphs (see \cite{VP11unlabeled} for details). Furthermore, the proof holds for other matrix norms (which might speed up convergence), and the regression scenario, where $|\mM|$ is infinite (again, see Methods for details).  
\end{thex}

% \subsection*{The practicality of a consistent statistical test for supervenience} % (fold)
